Run-Time Monitoring with Adjustable Overheads \\Using Dataflow-Guided Filtering

### 2014-09-13 dropping.tex
Various edits based on a full pass.

Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitor} performs a
series of \emph{monitoring operations}. The monitor operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events is sent to the monitor and
buffered in a FIFO structure to decouple the
running of the main core and the monitor. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitor may take several cycles to process a single event
from the main core. 
Monitors typically maintain a set of metadata corresponding to the main
program's memory. Similarly, monitors also typically maintain a set of metadata
corresponding to the main core's register file.
%We refer to these stalls and other overheads, such as
%contention for shared resources, as \emph{monitoring overheads}. 
Monitors typically maintain metadata corresponding to each memory location
({\tt mem\_metadata[addr]}) and register ({\tt rf\_metadata[reg\_id]}) of the
main core.
If the monitor
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

@@

Figure~\ref{fig:dropping.example_full} shows an example pseudo-code segment,
its assembly level instructions, and the corresponding monitoring operations
for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc} (line 1). 
As {\tt malloc} returns the array's address in a register, the monitor
associates base and bounds metadata with the corresponding metadata. 
Next, pointer {\tt y} is set to point to the middle of array {\tt x} (line 2). 
At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
The monitor propagates the metadata of the original pointer in {\tt r1} to the
metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
Line 3a shows setting register {\tt r3} to a constant value of 1.
When this happens,
{\tt r3}'s metadata is also reset to NULL in case it previously stored a
pointer. 
Finally, the value of {\tt r3} is written using both pointers {\tt x} and {\tt y}.
In both cases, the monitor checks whether the address that is written
to is within the bounds stored in the metadata. In the first case (line 3b), {\tt x+12}
is within the original array's bounds. No error is raised and the metadata of
{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
pointer, than this would allow a future instruction to load the pointer and use
it to access the corresponding array or data structure. In the second case,
{\tt y+12} (line 4) corresponds to {\tt x+20} which is not within the array's bounds and
the monitor will raise an error.

Note that this monitoring is performed automatically and transparently with
almost no modification needed to the main program. The only modifications to
the main program that are needed are to initially set metadata, such as setting
the base and bounds addresses on a {\tt malloc} call for array bounds check.
Otherwise, on the commit
of instruction types relevant to the specific monitor, the needed address and
register information is forwarded to the monitor automatically.
Here, we have shown the dynamic sequence of operations executed by the monitor but
the actual static code consists of a set of instructions to be run for each
possible instruction type (load, store, etc.).

@@

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as the check
for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false
positives is to add a 1-bit flag to each metadata in order to mark these
metadata as
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. In the example, no
monitoring operations are dropped. However, suppose that the monitoring for
line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

Figure~\ref{fig:dropping.example_null} shows an example of how this null filtering operates.
Here, the main core's code has been slightly modified from
Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
Without null metadata filtering, the instructions for lines 2 and 4 would be
forwarded to the monitor since the system does not know whether {\tt r2}
contains a pointer address or not. However, if we use a flag to mark {\tt r2}
as null when it is loaded with a constant, then we can propagate this null
information and filter out monitoring for lines 2 and 4.

The operations performed by null filtering are almost identical to the
operations needed for invalidation shown in
Figure~\ref{fig:dropping.example_invalid}. Instead of taking a logical OR of the
source invalid flags to determine whether monitoring can be skipped, null
metadata filtering takes a logical AND of the source null flags.
Thus, we can easily enable this null metadata filtering on our architecture by
extending the dataflow flags to be two bits wide. One bit is used to
keep track of invalidation information while the second bit is used to keep
track of null information. All flags are initialized to null and the filter
decision table is extended with the propagation and filtering decision rules
for null metadata filtering.
The results is a single hardware design that provides enables both partial
monitoring and null metadata filtering.

>>>>

Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitor} performs a
series of \emph{monitoring operations}. The monitor operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events is sent to the monitor and
buffered in a FIFO structure to decouple the
running of the main core and the monitor. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitor may take several cycles to process a single event
from the main core. 
Monitors typically maintain metadata corresponding to each memory location
({\tt mem\_metadata[addr]}) and register ({\tt rf\_metadata[reg\_id]}) of the
main core.
If the monitor
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

@@

Figure~\ref{fig:dropping.example_full} shows an example pseudo-code segment,
its assembly level instructions, and the corresponding monitoring operations
for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc} (line 1). 
As {\tt malloc} returns the array's address in a register, the monitor
associates base and bounds metadata with the corresponding register. 
Next, pointer {\tt y} is set to point to the middle of array {\tt x} (line 2). 
At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
The monitor propagates the metadata of the original pointer in {\tt r1} to the
metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
Line 3a shows setting register {\tt r3} to a constant value of 1.
When this happens,
{\tt r3}'s metadata is also reset to NULL in case it previously stored a
pointer. 
Finally, the value of {\tt r3} is written to memory using both pointers {\tt x} and {\tt y}.
In both cases, the monitor checks whether the store address is within the bounds of the register metadata. In the first case (line 3b), {\tt x+12}
is within the original array's bounds. No error is raised and the metadata of
{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
pointer, than this would allow a future instruction to load the pointer and use
it to access the corresponding array or data structure. In the second case,
{\tt y+12} (line 4) corresponds to {\tt x+20} which is not within the array's bounds and
the monitor will raise an error.

Note that this monitoring is performed automatically and transparently with
almost no modification needed to the main program. The only modifications to
the main program that are needed are to initially set metadata, such as setting
the base and bounds addresses on a {\tt malloc} call for array bounds check.
Later propagation and checks of the metadata occur automatically as
instructions are forwarded.
Here, we have shown the dynamic sequence of operations executed by the monitor but
the actual static code consists of a set of instructions to be run for each
possible instruction type (load, store, etc.). Only instruction types relevant
for the particular monitoring scheme are forwarded.

@@

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as the check
for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false
positives is to associate a 1-bit flag with each metadata in order to mark these
metadata as valid or
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. In the example, no
monitoring operations are dropped. However, suppose that the monitoring for
line 1 is dropped. When a monitoring event is dropped, metadata that would have
been updated is marked as invalid. In this case, {\tt rf\_invalid[1]} will be
marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

Figure~\ref{fig:dropping.example_null} shows an example of how this null filtering operates.
Here, the main core's code has been slightly modified from
Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
Without null metadata filtering, the instructions for lines 2 and 4 would be
forwarded to the monitor since the system does not know whether {\tt r2}
contains a pointer address or not. However, if we use a 1-bit flag to mark {\tt r2}
as null when it is loaded with a constant, then we can propagate this null
information and filter out monitoring for lines 2 and 4.

The operations performed by null filtering are almost identical to the
operations needed for invalidation shown in
Figure~\ref{fig:dropping.example_invalid}. Instead of taking a logical OR of the
source invalid flags to determine whether monitoring can be skipped, null
metadata filtering takes a logical AND of the source null flags.
Thus, we can easily enable this null metadata filtering on our architecture by
extending the dataflow flags to be two bits wide. One bit is used to
keep track of invalidation information while the second bit is used to keep
track of null information. All flags are initialized to null and the filter
decision table is extended with the propagation and filtering decision rules
for null metadata filtering.
The result is a single hardware design that enables both partial
monitoring and null metadata filtering.

### 2014-09-13 evaluation.tex
Various edits based on a full pass.

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We implement the monitor as a software-based monitor
running on a parallel processor core. We model the main and monitoring cores as
running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for flags.

In order to explore the generality of the architecture for
different monitors, we implemented five different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race detection (LS).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations. Additionally, recent
work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
LockSet attempts to detect possible race conditions in multi-threaded programs
by associating information about which locks have access to shared memory
locations. If a shared memory location is accessed without the use of a lock
then a race condition may exist. Since we implement the monitor using a
processor core and our dataflow engine was designed to be generally applicable,
the same hardware platform supports all of these monitoring techniques.

We tested our system using benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 2 billion instructions. 
% An initial slack of 2
% million cycles, which is less than 1\% of
% the total execution time, is given for a 1.1x target overhead. This initial
% slack is scaled proportionally for other overhead targets. 
Since LS race detection requires multi-threaded programs, we tested LS using
the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
200 million instructions.

@@

Figure~\ref{fig:evaluation.filtering} shows the
execution times of performing monitoring with null filtering enabled normalized
to the execution time of the benchmarks without monitoring. UMC sees normalized
execution times of 1.7-7x with an average of 4x.
For BC, normalized execution
are 3x on average and range from 1.1x to 8.3x.
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
DIFT sees an average of 10\% overhead with a maximum overhead of 31\% with null
metadata filtering. This is due, in part, to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would higher overheads. IMP shows normalized
executions from 2-6x with an average of 4x. Overheads for LS (not pictured) are
1.03-1.72x after null metadata filtering and the average
overhead observed is 1.33x for the SPLASH-2 benchmarks.
These overheads are similar to the overheads seen by previous multi-core
monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
\cite{fade-hpca14} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

In this section, we evaluate the effectiveness of
using partial monitoring to trade-off coverage for reduced overheads.
For these results, we use a slack-based, unrestricted dropping policy.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of dynamic checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system, 
assuming that errors/attacks are uniformly distributed across checks. 
%While we could not evaluate real errors/attacks due to difficulty in setting up
%a large number of bugs and exploits, we believe that the percentage of checks
%provides a good estimate of detection probability when errors/attacks are 
%uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

We see that by varying the overhead budget, the coverage achieved also varies.
With only a 1.1x overhead target (10\% slowdown), array bounds check still
achieves 84\% monitoring coverage on average. The high coverage
achieved with such low overheads is due to two main effects.  The first is that
monitoring can be done in parallel, providing a certain level of monitoring coverage without
introducing overheads. The second effect is that, although we aggressively
filter out monitoring events, there may still exist a large number of
monitoring events that do not lead to security or reliability checks. As a
result, dropping these events can reduce overheads without a large impact on
monitoring coverage.

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.
Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
71\% coverage with a 1.2x overhead. 
Since the overheads of DIFT are low to begin with, setting an overhead target
1.1x is enough for all benchmarks tested to reach 100\% coverage.
\DAN{Try to explain water\_nsquared results not reaching 100\%?}

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We compare these percentages against the case when full monitoring
is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the min, max, and average error across
instruction types for each benchmark and for varying overheads. We see that
with a 1.1x overhead a max error of 9\% is observed across benchmarks and the
average error across instruction types and benchmarks is less than 2\%.

@@

In addition to evaluating our system for a core-based monitor, we also
evaluated partial monitoring on a higher performance FPGA-based monitor. This
setup is based on the FlexCore \cite{flexcore-micro10} setup and uses a
fully-pipelined monitor running on an FPGA fabric at half the frequency of the
main core. We show results for UMC due to space constraints though other
monitors show similar trends. In this case, the full overheads of
monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
the coverage achieved as we sweep the overhead target from 1.01-1.7x. We see that partial
monitoring can allow the overheads of such a system to be pushed below 10\%
while still providing over 50\% coverage.

@@

From these results, we see that for monitors which many short, independent
chains, such as UMC, source-only dropping can provide better coverage than
unrestricted dropping with the same overhead. For other monitors, such as BC,
source-only dropping is prone to overshoot the overhead target.
Thus, for applications where staying with an overhead target is especially
important, such as soft real-time systems, a slack-based, unrestricted dropping
policy is more appropriate. On the other hand, if occasional slowdowns are
acceptable, then source-only dropping can provide better coverage on average.

@@

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored in at
least one of the runs. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that it is likely that at least one
monitoring event in a chain will be dropped with the unrestricted dropping
policy. 
Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.

@@

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 1.5x monitoring overhead target. The peak power is 41-72 mW, which is
less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43
mW, corresponding to 5-11\% of the main core's runtime power.

>>>>

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We implement the monitor as a software-based monitor
running on a parallel processor core. We model the main and monitoring cores as
running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.

In order to explore the generality of the architecture for
different monitors, we implemented five different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race detection (LS).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
described in Section~\ref{sec:dropping}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations or to detect malicious
software \cite{tang-raid14}.
LockSet attempts to detect possible race conditions in multi-threaded programs
by tracking metadata about which locks are protecting shared memory
locations. If a shared memory location is accessed while unprotected
then a race condition may exist. Since we implement the monitor using a
processor core and our dataflow engine was designed to be generally applicable,
the same hardware platform supports all of these monitoring techniques.

We tested our system using benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown.
We simulated each benchmark for 2 billion instructions. 
% An initial slack of 2
% million cycles, which is less than 1\% of
% the total execution time, is given for a 1.1x target overhead. This initial
% slack is scaled proportionally for other overhead targets.

Since LockSet race detection requires multi-threaded programs, we tested LS using
the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
200 million instructions with 2 main cores running the benchmark application.
Each main core was connected to a dataflow engine and a monitoring core.

@@

Figure~\ref{fig:evaluation.filtering} shows the
execution times of performing monitoring with null filtering enabled normalized
to the execution time of the benchmarks without monitoring. UMC sees normalized
execution times of 1.7-7x with an average of 4x.
For BC, normalized execution
are 3x on average and range from 1.1x to 8.3x.
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
DIFT sees an average of 1.1x slowdown with a maximum slowdown of 1.3x with null
metadata filtering. This low overhead is due, in part, to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would higher overheads. IMP shows normalized
execution times of 2-6x with an average of 4x. Overheads for LS (not pictured) are
1.03-1.72x after null metadata filtering and the average
overhead observed is 1.33x for the SPLASH-2 benchmarks.
Our baseline system and overheads are similar previous multi-core
monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
\cite{fade-hpca14} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we also evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

In this section, we evaluate the effectiveness of
using partial monitoring to trade-off coverage for reduced overheads.
For these results, we use a slack-based, unrestricted dropping policy.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of dynamic checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC and LS). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system, 
assuming that errors/attacks are uniformly distributed across checks. 
%While we could not evaluate real errors/attacks due to difficulty in setting up
%a large number of bugs and exploits, we believe that the percentage of checks
%provides a good estimate of detection probability when errors/attacks are 
%uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

We see that by varying the overhead budget, the coverage achieved also varies.
With only a 1.1x overhead target (10\% slowdown), array bounds check still
achieves 84\% monitoring coverage on average. The high coverage
achieved with such low overheads is due to two main effects.  The first is that
monitoring can be done in parallel, providing a certain level of monitoring coverage without
introducing overheads. The second effect is that there may still exist a large number of
monitoring events that do not lead to security or reliability checks. As a
result, dropping these events can reduce overheads without a large impact on
monitoring coverage.

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.
Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
71\% coverage with a 1.2x overhead. 
Since the overheads of DIFT are low to begin with, setting an overhead target
1.1x is enough for all benchmarks tested to reach 100\% coverage.

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We take the difference of these percentages compared to the case
when full monitoring
is performed to calculate the error. Figure~\ref{fig:evaluation.imp_sweep}
shows the min, max, and average error across
instruction types for each benchmark and for varying overheads. We see that
with a 1.1x overhead a max error of 9\% is observed across benchmarks and the
average error across instruction types and benchmarks is less than 2\%.

@@

In addition to evaluating our system for a core-based monitor, we also
evaluated partial monitoring on a higher performance FPGA-based monitor. This
setup is based on the FlexCore \cite{flexcore-micro10} setup and uses a
fully-pipelined monitor running on an FPGA fabric at half the frequency of the
main core. We only show results for UMC due to space constraints though other
monitors show similar trends. In this case, the full overheads of
monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
the coverage achieved as we sweep the overhead target from 1.01-1.7x. We see that partial
monitoring can allow the overheads of such a system to be pushed below 10\%
while still providing over 50\% coverage.

@@

From these results, we see that for monitors which many short, independent
chains, such as UMC, source-only dropping can provide better coverage than
unrestricted dropping with the same overhead. For other monitors, such as BC,
source-only dropping is prone to overshoot the overhead target.
Thus, for applications where staying with an overhead target is especially
important, such as soft real-time systems, a slack-based, unrestricted dropping
policy is more appropriate. On the other hand, if maximum coverage is desired
and occasional slowdowns are acceptable, then source-only dropping can provide
better coverage on average.

@@

One application of partial monitoring with low overheads is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored in at
least one of the runs. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that it is likely that at least one
monitoring event in a chain will be dropped with the unrestricted dropping
policy. 
Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.

@@

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks running with
a 1.5x monitoring overhead target. The peak power is 31-72 mW, which is
less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 13-43
mW, corresponding to 4-11\% of the main core's runtime power.

### 2014-09-13 introduction.tex
Various edits based on a full pass.

% Adjustable overhead
Partial monitoring with adjustable overheads is achieved by dynamically
dropping monitoring operations when overheads exceed a specified overhead
budget. Although this results in reduced monitoring coverage (i.e., false
negatives), false positives can also occur due to out-of-date metadata information.
In order to prevent these
false positive, we need to track the dropped metadata flows.  We present a
hardware-based dataflow tracking engine that keeps track of these dropped
flows. With this architecture, we investigate different policies for deciding
which events to drop for partial monitoring. These different policies show a
trade-off between closely matching the overhead budget and the monitoring
coverage achieved.

% Filtering
Recent studies have shown that filtering out monitoring operations on clean or null
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
For example, for an array bounds check, checking non-pointer accesses is
unnecessary. In addition to enabling partial monitoring, we show how a simple
extension to our dataflow engine enables this null metadata filtering.

The main contributions of this paper are summarized as follows:
\begin{itemize}
  \item We present a hardware architecture using a dataflow tracking engine to
  efficiently enable partial monitoring while preventing false positives. This
  hardware architecture is designed to be generally applicable to a wide-range
  of run-time monitoring techniques.
  \item We investigate multiple policies on when and which operations to drop 
  for partial monitoring, and show the trade-offs in the design space.
  \item We extend the dataflow engine to enable filtering null metadata to
  reduce overheads.
\end{itemize}

@@

% Evaluation
In order to evaluate our approach, we applied it to 4 different monitoring
techniques. These monitoring techniques vary in what events they monitor, the
size and semantics of their metadata, and the operations performed on metadata.
% Our experiments show that filtering reduces the average overhead of these
% monitors from 7-18x down to 13-300\%.
% In addition, with partial monitoring, for the monitoring schemes which still
% see average overheads of 300\%, we show that with just 10\% overheads, we can still
% provide 32-82\% of coverage on average. By increasing the
% overhead budget, this coverage rate can be increased. 
These monitors show slowdowns of 1.1-4.3x with the null metadata filtering when
implemented on a multi-core platform.
%4-16x without null metadata filtering
%and 1.1-4.3x after null metadata filtering.  
Our results show that the partial monitoring can still achieve coverage of 43-82\%
on average at the reduced slowdown of 1.5x. 
The results also show that overhead and coverage can be varied to create a trade-off.

>>>>

% Adjustable overhead
Partial monitoring with adjustable overheads is achieved by dynamically
dropping monitoring operations when overheads exceed a specified overhead
budget. Although this results in reduced monitoring coverage (i.e., false
negatives), false positives can also occur due to out-of-date metadata information.
In order to prevent these
false positive, we need to track the dropped metadata flows.  We present a
hardware-based dataflow tracking engine that keeps track of these dropped
flows. With this architecture, we investigate different policies for deciding
which events to drop for partial monitoring. These different policies show a
trade-off between closely matching the overhead budget and the monitoring
coverage achieved.
In addition to enabling partial monitoring, we show how a simple
extension to our dataflow engine enables this null metadata filtering.
For example, for an array bounds check, checking non-pointer accesses is
unnecessary and can be filtered out using the dataflow engine.

The main contributions of this paper are summarized as follows:
\begin{itemize}
  \item We present a hardware architecture using a dataflow tracking engine to
  efficiently enable partial monitoring while preventing false positives. This
  hardware architecture is designed to be generally applicable to a wide-range
  of run-time monitoring techniques and monitor implementations.
  \item We investigate multiple policies on when and which operations to drop 
  for partial monitoring, and show the trade-offs in the design space.
  \item We extend the dataflow engine to enable filtering null metadata to
  reduce overheads.
\end{itemize}

@@

% Evaluation
In order to evaluate our approach, we applied it to five different monitoring
techniques. These monitoring techniques vary in what events they monitor, the
size and semantics of their metadata, and the operations performed on metadata.
% Our experiments show that filtering reduces the average overhead of these
% monitors from 7-18x down to 13-300\%.
% In addition, with partial monitoring, for the monitoring schemes which still
% see average overheads of 300\%, we show that with just 10\% overheads, we can still
% provide 32-82\% of coverage on average. By increasing the
% overhead budget, this coverage rate can be increased. 
These monitors show slowdowns of 1.1-4.3x with the null metadata filtering when
implemented on a multi-core platform.
Our results show that the partial monitoring can still achieve significant
coverage with reduced overheads. For example, for monitors which showed 3-4x
average overheads, slowdown could be reduced to 1.5x while still achieving a
coverage of 43-82\% on average.
